---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
1. Download the csv file activity.csv into your pwd.

2. Read the csv file. Hint : Use the read.csv() function
* code  for reading in the dataset
* 
```{r}
MyData <- read.csv(file="E://R//course5//Course5//activity.csv", header=TRUE, sep=",")
str(MyData)
```

* preprocessing the data



## What is mean total number of steps taken per day?
```{r}
avgsteps <- mean(MyData$Steps)
str(avgsteps)
```
average number of steps taken by this person is `r avgsteps` .

*histogram of the total number of steps taken each day


## What is the average daily activity pattern?
*Both the mean and median are reported
*the 5-minute interval containing the maximum number of steps is reported

```{r}
average_steps_day <- aggregate(MyData$Steps,list(Day=MyData$Day),mean )
dailyaverage <-aggregate(Steps ~ Day, MyData,mean)
average_steps_day
colnames(average_steps_day) <- c("day","steps")
head(average_steps_day)
```

```{r}
medianofstepsbyday= median(average_steps_day$steps)
medianofstepsbyday
```
*the mean and median are reported
+ the mean  : `r average_steps_day` 
+ the median: `r medianofstepsbyday` 


## Imputing missing values
* description of a strategy for imputing missing data

* a histogram of the total number of steps taken each day after missing values were impute


## Are there differences in activity patterns between weekdays and weekends?
*a panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
dailyaverage2<-aggregate(Steps ~ Hours, myData2,mean)

ggplot(dailyaverage2,aes(x=hours,y=Steps)) +
  geom_line(color="blue",size=1)+
labs(title="average pattern",x="hours",y="Steps")

```

*  report show all of the R code needed to reproduce the results (numbers, plots, etc.) in the report
* this report contains R code for **all** of the results presented

